---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/pachadotdev/canadamaps/workflows/R-CMD-check/badge.svg)](https://github.com/pachadotdev/canadamaps/actions)
<!-- badges: end -->
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# canadamaps

## General idea

The idea is to avoid "duplications", for example instead of adding a provinces 
map or others, we provide functions to sum Census Divisions in all possible 
cases.

```{r general, message=FALSE, warning=FALSE}
library(ggplot2)
library(canadamaps)

ggplot(data = census_divisions) + 
  geom_sf(aes(geometry = geometry)) +
  labs(title = "Canada's Census Divisions")

# same idea as the same plot, different aggregation
#
# ggplot(data = get_agricultural_divisions()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Census Agricultural Regions")
# 
# ggplot(data = get_economic_regions()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Economic Regions")
# 
# ggplot(data = federal_electoral_districts) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Federal Electoral Districts")
# 
# ggplot(data = get_provinces()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Provinces")
```

## Using real data

We start by loading the required packages.

```{r pkgs, warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(sf)
```

Let's say I want to replicate the map from [Health Canada](https://health-infobase.canada.ca/covid-19/vaccination-coverage/), which was checked on 2023-08-02 and was updated up to 2023-06-18. To do this, I need to download the [CSV file](https://health-infobase.canada.ca/src/data/covidLive/vaccination-coverage-map.csv) from Health Canada and then combine it with the provinces map from canadamaps.

```{r data, warning=FALSE, message=FALSE}
url <- "https://health-infobase.canada.ca/src/data/covidLive/vaccination-coverage-map.csv"
csv <- paste0("data_processing/", gsub(".*/", "", url))
if (!file.exists(csv)) download.file(url, csv)

vaccination <- read_csv(csv) %>% 
  filter(week_end == as.Date("2023-06-18"), pruid != 1) %>% 
  select(pruid, proptotal_atleast1dose)

vaccination <- vaccination %>% 
  left_join(get_provinces(), by = "pruid") %>% # canadamaps in action
  mutate(
    label = paste(gsub(" /.*", "", prname),
                  paste0(proptotal_atleast1dose, "%"), sep = "\n"),
  )
```

An initial plot can be done with the following code.

```{r plot, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}
# colours obtained with Chromium's inspector
colours <- c("#efefa2", "#c2e699", "#78c679", "#31a354", "#006837")

vaccination %>% 
  ggplot() +
  geom_sf(aes(fill = proptotal_atleast1dose, geometry = geometry)) +
  geom_sf_label(aes(label = label, geometry = geometry)) +
  scale_fill_gradientn(colours = colours, name = "Cumulative percent") +
  labs(title = "Cumulative percent of the population who have received at least 1 dose of a COVID-19 vaccine") +
  theme_minimal(base_size = 13)
```

What if we want a Lambert (conic) projection? We can change the CRS with the sf package but please read the explanation from [Stats Canada](https://www150.statcan.gc.ca/n1/pub/92-195-x/2011001/other-autre/mapproj-projcarte/m-c-eng.htm).

```{r lambert, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}
vaccination$geometry <- st_transform(vaccination$geometry,
  crs = "+proj=lcc +lat_1=49 +lat_2=77 +lon_0=-91.52 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs")

vaccination %>% 
  ggplot() +
  geom_sf(aes(fill = proptotal_atleast1dose, geometry = geometry)) +
  geom_sf_label(aes(label = label, geometry = geometry)) +
  scale_fill_gradientn(colours = colours, name = "Cumulative percent") +
  labs(title = "Cumulative percent of the population who have received at least 1 dose of a COVID-19 vaccine") +
  theme_minimal(base_size = 13)
```

Finally, we can use a different ggplot theme.

```{r lambert2, warning=FALSE, message=FALSE, fig.width=10, fig.height=6}
vaccination %>% 
  ggplot() +
  geom_sf(aes(fill = proptotal_atleast1dose, geometry = geometry)) +
  geom_sf_label(aes(label = label, geometry = geometry)) +
  scale_fill_gradientn(colours = colours, name = "Cumulative percent") +
  labs(title = "Cumulative percent of the population who have received at least 1 dose of a COVID-19 vaccine") +
  theme_void() +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)
  )
```

## Units of aggregation

### Census Division

The finest division in this package is the Census Division (CDs) which can be
of the next types (reference: https://www.statcan.gc.ca/en/subjects/standard/sgc/2011/sgc-tab-d).

|Language form of CD type|Abbreviation for English language publications|Title for English language publications|Abbreviation for French language publications|Title for French language publications|Abbreviation for bilingual publications|Title for bilingual publications         |
|------------------------|----------------------------------------------|---------------------------------------|---------------------------------------------|--------------------------------------|---------------------------------------|-----------------------------------------|
|Bilingual               |CDR                                           |Census division                        |CDR                                          |Division de recensement               |CDR                                    |Census division / Division de recensement|
|Bilingual               |CT                                            |County                                 |CT                                           |Comté                                 |CT                                     |County / Comté                           |
|English only            |CTY                                           |County                                 |CTY                                          |County                                |CTY                                    |County                                   |
|English only            |DIS                                           |District                               |DIS                                          |District                              |DIS                                    |District                                 |
|English only            |DM                                            |District municipality                  |DM                                           |District municipality                 |DM                                     |District municipality                    |
|French only             |MRC                                           |Municipalité régionale de comté        |MRC                                          |Municipalité régionale de comté       |MRC                                    |Municipalité régionale de comté          |
|English only            |RD                                            |Regional district                      |RD                                           |Regional district                     |RD                                     |Regional district                        |
|English only            |REG                                           |Region                                 |REG                                          |Region                                |REG                                    |Region                                   |
|English only            |RM                                            |Regional municipality                  |RM                                           |Regional municipality                 |RM                                     |Regional municipality                    |
|French only             |TÉ                                            |Territoire équivalent                  |TÉ                                           |Territoire équivalent                 |TÉ                                     |Territoire équivalent                    |
|Bilingual               |TER                                           |Territory                              |TER                                          |Territoire                            |TER                                    |Territory / Territoire                   |
|English only            |UC                                            |United counties                        |UC                                           |United counties                       |UC                                     |United counties                          |

The division type is specified in the `census_divisions` table.

### Census Agricultural Regions

Census Agricultural Regions (CARs) can be obtained as sums of CDs. Excluding
some special cases for Northwestern Territories, Nunavur and Yukon that we
clarified over email communication, the source to match CDs to CARs was obtained
from [Census of Agriculture Reference Maps](https://www150.statcan.gc.ca/n1/pub/95-630-x/95-630-x2017000-eng.htm)
and manually organized in a spreadsheet (https://github.com/pachadotdev/canadamaps/tree/main/data_xlsx).

### Economic Regions

Economic Regions (ERs) can be obtained as sums of CDs. The only special case is
the Halton, which belongs to two economic zones and it's the only CD that has
to be carefully separated (i.e. see
https://github.com/pachadotdev/canadamaps/blob/main/data_processing/02_census_divisions_and_derivatives.R).

### Federal Electoral Districts

These cannot be obtained as sums of CDs, therefore these are stored in their
own table `federal_electoral_districts`.
