---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/pachadotdev/canadamaps/workflows/R-CMD-check/badge.svg)](https://github.com/pachadotdev/canadamaps/actions)
<!-- badges: end -->
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# canadamaps

## General idea

The idea is to avoid "duplications", for example instead of adding a provinces 
map or others, we provide functions to sum Census Divisions in all possible 
cases.

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(canadamaps)

ggplot(data = census_divisions) + 
  geom_sf(aes(geometry = geometry)) +
  labs(title = "Canada's Census Divisions")

# same idea as the same plot, different aggregation
#
# ggplot(data = get_agricultural_divisions()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Census Agricultural Regions")
# 
# ggplot(data = get_economic_regions()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Economic Regions")
# 
# ggplot(data = federal_electoral_districts) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Federal Electoral Districts")
# 
# ggplot(data = get_provinces()) + 
#   geom_sf(aes(geometry = geometry)) +
#   labs(title = "Canada's Provinces")
```

## Using real data

Example (after loading the packages above): Let's say I want to replicate
the map from https://health-infobase.canada.ca/covid-19/vaccination-coverage/, which was
checked on 2021-10-23 and was updated up to 2021-10-16.

```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)

url <- "https://health-infobase.canada.ca/src/data/covidLive/vaccination-coverage-map.csv"
csv <- paste0("data_processing/", gsub(".*/", "", url))
if (!file.exists(csv)) download.file(url, csv)

colours <- c("#efefa2", "#c2e699", "#78c679", "#31a354", "#006837")

vaccination <- read_csv(csv) %>% 
  filter(week_end == as.Date("2021-10-16"), pruid != 1) %>% 
  select(pruid, proptotal_atleast1dose)

vaccination <- vaccination %>% 
  left_join(get_provinces(), by = "pruid") %>% # canadamaps in action
  mutate(
    label = paste(gsub(" /.*", "", prname),
                  paste0(proptotal_atleast1dose, "%"), sep = "\n"),
  )

# devtools::install_github("yutannihilation/ggsflabel")

vaccination %>% 
  ggplot() +
  geom_sf(aes(fill = proptotal_atleast1dose, geometry = geometry)) +
  # geom_sf_label(aes(label = label, geometry = geometry)) +
  # improve label positioning with
  ggsflabel::geom_sf_label_repel(aes(label = label, geometry = geometry)) +
  scale_fill_gradientn(colours = colours, name = "Cumulative percent") +
  labs(title = "Cumulative percent of the population who have received at least 1 dose of a COVID-19 vaccine") +
  theme_minimal(base_size = 13)
```

## Units of aggregation

### Census Division

The finest division in this package is the Census Division (CDs) which can be
of the next types (reference: https://www.statcan.gc.ca/en/subjects/standard/sgc/2011/sgc-tab-d).

|Language form of CD type|Abbreviation for English language publications|Title for English language publications|Abbreviation for French language publications|Title for French language publications|Abbreviation for bilingual publications|Title for bilingual publications         |
|------------------------|----------------------------------------------|---------------------------------------|---------------------------------------------|--------------------------------------|---------------------------------------|-----------------------------------------|
|Bilingual               |CDR                                           |Census division                        |CDR                                          |Division de recensement               |CDR                                    |Census division / Division de recensement|
|Bilingual               |CT                                            |County                                 |CT                                           |Comté                                 |CT                                     |County / Comté                           |
|English only            |CTY                                           |County                                 |CTY                                          |County                                |CTY                                    |County                                   |
|English only            |DIS                                           |District                               |DIS                                          |District                              |DIS                                    |District                                 |
|English only            |DM                                            |District municipality                  |DM                                           |District municipality                 |DM                                     |District municipality                    |
|French only             |MRC                                           |Municipalité régionale de comté        |MRC                                          |Municipalité régionale de comté       |MRC                                    |Municipalité régionale de comté          |
|English only            |RD                                            |Regional district                      |RD                                           |Regional district                     |RD                                     |Regional district                        |
|English only            |REG                                           |Region                                 |REG                                          |Region                                |REG                                    |Region                                   |
|English only            |RM                                            |Regional municipality                  |RM                                           |Regional municipality                 |RM                                     |Regional municipality                    |
|French only             |TÉ                                            |Territoire équivalent                  |TÉ                                           |Territoire équivalent                 |TÉ                                     |Territoire équivalent                    |
|Bilingual               |TER                                           |Territory                              |TER                                          |Territoire                            |TER                                    |Territory / Territoire                   |
|English only            |UC                                            |United counties                        |UC                                           |United counties                       |UC                                     |United counties                          |

The division type is specified in the `census_divisions` table.

### Census Agricultural Regions

Census Agricultural Regions (CARs) can be obtained as sums of CDs. Excluding
some special cases for Northwestern Territories, Nunavur and Yukon that we
clarified over email communication, the source to match CDs to CARs was obtained
from [Census of Agriculture Reference Maps](https://www150.statcan.gc.ca/n1/pub/95-630-x/95-630-x2017000-eng.htm)
and manually organized in a spreadsheet (https://github.com/pachadotdev/canadamaps/tree/main/data_xlsx).

### Economic Regions

Economic Regions (ERs) can be obtained as sums of CDs. The only special case is
the Halton, which belongs to two economic zones and it's the only CD that has
to be carefully separated (i.e. see
https://github.com/pachadotdev/canadamaps/blob/main/data_processing/02_census_divisions_and_derivatives.R).

### Federal Electoral Districts

These cannot be obtained as sums of CDs, therefore these are stored in their
own table `federal_electoral_districts`.
